# Practical-Machine-Learning---Final-Project

---
title: "FinalProject"
output: html_document
---
The goal of your project is to predict the manner in which they did the exercise.
This is the "classe" variable in the training set. 

#Load data
```{r}
train = read.csv("pml-training.csv")
test = read.csv("pml-testing.csv")

str(train)
summary(train$classe)
```


#Dealing with NA and missing data.  Convert blank and DIV/0 to NA.  
#Remove variables with NA data
```{r}
training <- train
training[training == ""] <- NA
training[training == "#DIV/0!"] <- NA

hasMissing <- unlist(lapply(training, function(x) any(is.na(x))))
hasMissing <- names(hasMissing)[hasMissing]
hasMissing

`%ni%` <- Negate(`%in%`)
training <- subset(training,select = names(training) %ni% hasMissing)
str(training)
training <- select(training, -(1:7))
training <- select(training, -c(classe))  # remove classe
str(training)
summary(training)

```


#Check for possible transformations: skewness, near zero variance
```{r}

library(e1071)
skewValues <- apply(training, 2, skewness) # 1 indicates rows, 2 indicates columns
skewValues

nzv = nearZeroVar(training, saveMetrics = TRUE)

```


# administer a series of transformation
```{r}
trans <- preProcess(training, method = c("BoxCox", "center", "scale", "pca" ))
trans
pcaObject <- prcomp(training, center = TRUE, scale. = TRUE)
```

# apply the transformation
```{r}
transformed <- predict(trans, training)
head(transformed[ , 1:5])
```

# add class variable back to data
```{r}
transformed$classe <- train$classe
```

# RF with PCA (narrows down to 25 variables)
```{r}
rfmodel = train(classe ~ . , data = transformed, method = "rf",model=FALSE)
```

# save a copy of model
```{r}
saveRDS(rfmodel, file="rfmodel.rds")
rfmodel = readRDS("rfmodel.rds")
```

# Check Accuracy : 0.983
```{r}
rfpredict = predict(rfmodel$final, data = transformed)
rfaccuracy = confusionMatrix(rfpredict, transformed$classe)
```

# model on test data
# Pre-processing and apply transformation on test set
```{r}
testing <- test
`%ni%` <- Negate(`%in%`)
testing <- subset(testing,select = names(testing) %ni% hasMissing)
str(testing)
testing <- select(testing, -(1:7))
testing <- select(testing, -c(problem_id))

transformed_test <- predict(trans, testing) 
head(transformed_test[ , 1:5])

rfpredict_testing = predict(rfmodel$final, newdata = transformed_test)
rfpredict_testing
```

# create txt files for submitting answers
```{r}
answers = rfpredict_testing
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)

```

